---
title: "PA1_template"
author: "Olena"
date: "14 12 2020"
output: html_document
---

## This is my variant of solving assigment 1 of Coursera course "Reproducible Research".
# What is mean total number of steps taken per day?
## The histogram shows the total number of steps per day.
```{r, echo=TRUE}
act <- read.csv(file = "activity.csv")
stepsday <- aggregate(steps ~ date, act, sum)
hist(stepsday$steps, col = "red", main = "Histogram of total number of steps per day", xlab = "Total number of steps per day")
```

## Mean of total number of steps per day is 10766, median is 10765.
```{r, echo=TRUE}
summary(stepsday)
```

# What is the average daily activity pattern?
## The plot shows the average number of steps over all day.
```{r, echo=TRUE}
stepsinterval <- aggregate(steps ~ interval, act, mean)
plot(stepsinterval$interval, stepsinterval$steps, type = "l", main = "Average number of steps over all day", xlab = "Interval", ylab = "Average number of steps")
```

## Which interval contains the maximum number of steps?
```{r, echo=TRUE}
maxsteps <- which.max(stepsinterval$steps)
stepsinterval[maxsteps, ]
```

# Imputing missing values
## The total number of missing value is:
```{r, echo=TRUE}
sum(is.na(act))
```

## The new dataset is equal to the original, but filled in by replacing NA's with the mean of interval.
```{r, echo=TRUE}
newact <- act
for (i in 1:nrow(newact)) {
  if(is.na(newact$steps[i])){
    newinterwal <- newact$interval[i]
    newsteps <- stepsinterval[
      stepsinterval$interval == newinterwal,]
    newact$steps[i] <- newsteps$steps
  }
}
```

## The new histogram (imputed NA's) shows the total number of steps per day.
```{r, echo=TRUE}
newstesday <- aggregate(steps ~ date, newact, sum)
hist(newstesday$steps, col = "green", main = "Histogram of total number of steps per day (imputed NA's)", xlab = "Total number of steps per day")
```

## Mean of total number of steps per day is 10766, new median is 10766.
```{r, echo=TRUE}
summary(newstesday)
```

# Are there are differences in activity patterns between weekdays and weekends?
## Create a new factor variables in the dataset - "weekday" and "weekend"
```{r, echo=TRUE}
Sys.setlocale("LC_TIME", "English")
newact$date <- as.Date(newact$date)
newact$day <- weekdays(newact$date)
weekday <- newact[(!newact$day %in% c("Saturday", "Sunday")), ]
weekend <- newact[(newact$day %in% c("Saturday", "Sunday")), ]
wdaysteps <- weekday[,1]
m <- matrix(wdaysteps, nrow = 288)
weekdaymean <- apply(m, 1, mean)
wendsteps <- weekend[,1]
m <- matrix(wendsteps, nrow = 288)
weekendmean <- apply(m, 1, mean)
```

## Plots show differences in activity patterns between weekdays and weekends
```{r, echo=TRUE}
par(mfrow=c(2,1))
plot(newact$interval[1:288], weekdaymean, type = "l", xlab = "Intervals", ylab = "Number of steps", col = "blue", lwd=2, main = "Weekday")
plot(newact$interval[1:288], weekendmean, type = "l", xlab = "Intervals", ylab = "Number of steps", col = "red", lwd=2, main = "Weekend")
```


